Cosmology with Weak Lensing photometric surveys: constraints with redshift tomography of non--Gaussian statistics

### 2016-04-05 photoz_paper/paper.tex
fix typo
implemented Zoltan edits 1

\title{Cosmology with photometric surveys: constraints with redshift tomography of non--Gaussian statistics}

@@

\section{Introduction}
%
Weak Gravitational Lensing is a promising technique to probe the Large Scale Structure of the universe in which the tracers are unbiased \citep{wlreview}. This technique has the potential of improving significantly the constraints on the Dark Energy equation of state $w$ because it is most sensitive to the matter density fluctuations at the non--linear stage. Cosmology inferences from Weak Lensing observations have been produced for past (CFHTLenS \citep{cfht1}) and current (DES \citep{DES}) surveys, and are being planned for future experiments as well (LSST \citep{LSST}, WFIRST \citep{WFIRST}, Euclid \citep{Euclid}). Because of the non-linear nature of the density fluctuations probed by Weak Lensing, cosmological information might leak from quadratic statistics (such as two point functions and power spectra) into more complicated non-Gaussian statistics, for which forward modeling from cosmological parameters requires numerical simulations of cosmic shear fields. An extensive list of these non-Gaussian statistics, and their cosmological information content, have been investigated extensively in the literature. We investigate the constraining power of a subset of these non-Gaussian statistics combined with redshift tomography in a LSST-like survey. We also investigate the effects of uncorrected photometric redshift systematics on parameter constraints when using redshift tomography. This work is organized as follows: we give an outline of the shear simulations we use in this work, followed by descriptions of the convergence reconstruction procedure and forward modeling of galaxy shape and photometric redshift systematics. We then give an overview of the parameter inferences techniques we used to forecast constraints on cosmology. We then discuss the main results of this work and present our conclusions as well as prospects for future work.

@@

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating mock shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses until $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the lensing ODE up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from Large Scale Structure, assuming the non lensed shape is a circle. 
We simulate $N_g = 10^6$ random galaxy positions $\{\pmb{\theta}_g\}$ distributed uniformly in a field of view of size $\theta_{\rm FOV}^2=(3.5{\rm deg})^2$, which correspond to a galaxy angular density of $n_g=22\,\rm{arcmin}^{-2}$. The galaxy have a distribution in redshift which mimics the one expected in the LSST survey
\begin{equation}
\label{meth:galdistr}
n(z) = n_0\left(\frac{z}{z_0}\right)^2\exp\left(-\frac{z}{z_0}\right)
\end{equation}  
%
with $z_0=0.3$ and $n_0$ a normalization constant fixed so that $n(z)$ integrates to the total number of galaxies $N_g$. For each galaxy we compute the cosmic shear at $\pmb{\theta}_g$ using equation (\ref{meth:sheareqn}), producing a shear catalog $\{\pmb{\gamma}_g\}$. Different random realizations of a shear catalog $\{\pmb{\gamma}_g\}_r$ can be obtained rotating and periodically shifting the Large Scale Structure in the $N$--body snapshots according to the procedure explained in \citep{PetriVariance}. We produce $N_r=16000$ pseudo--independent realizations of the shear catalog $\{\pmb{\gamma}_g\}$. These shear realization cover the total survey area of LSST 10 times.
We repeat the above procedure for $P=100$ different combinations of the parameter triplet $\bb{p}$, sampled according to a Latin hypercube scheme. The sampling procedure is the same described in \citep{CFHTMink,CFHTPeaks}. The parameter space sampling we adopted for our simulations is shown in Figure \ref{fig:simdesign}

\begin{figure*}
\includegraphics[scale=0.4]{Figures/design.eps}
\caption{Parameter space sampling chosen for our simulations. We show both the sampling in the $(\Omega_m,w)$ (left) and $(\Omega_m,\sigma_8)$ (right) slices of the parameter space. The fiducial cosmology has been marked in red.}
\label{fig:simdesign}
\end{figure*}
%
For each of the parameter choices in Figure \ref{fig:simdesign}, the $N$--body initial conditions are generated using the same random seed. In addition to this simulations, we produce simulated shear catalogs for a fiducial $\Lambda$CDM universe with $\bb{p}_0=(0.26,-1,0.8)$. In this case the randomization procedure is based on 5 independent $N$--body simulations, and the same number of pseudo--independent catalog realizations is produced.

@@

\subsection{Forward modeling of systematics}
We give an overview of the shear systematics included in this work. The measured galaxy ellipticity $\pmb{\epsilon}$ is an estimate of the cosmic shear $\pmb{\gamma}$ due to Large Scale Structure if the non--lensed galaxy shape is a circle. If this assumption is relaxed, the measured galaxy ellipticity $\pmb{\epsilon}_{\rm m}$ can be modeled as a cosmic shear term plus a noise term \citep{wlreview}
\begin{equation}
\pmb{\epsilon}_{\rm m} = \pmb{\gamma} + \pmb{\epsilon}_{\rm n}
\end{equation} 
%
where $\pmb{\epsilon}_{\rm n}$ is a random Gaussian variable with zero mean and redshift dependent variance $\sigma_n(z)=0.15+0.035z$. This is equivalent to say that the cosmic shear inferred from ellipticity observations $\pmb{\gamma}_{\rm m}$ can be written as the sum of the true cosmic shear plus a noise term $\pmb{\gamma}_{\rm n}$ with the same statistical properties as $\pmb{\epsilon}_{\rm n}$. 
We add independent random realizations of the shape noise $\pmb{\gamma}_{\rm n}$ to each of the $N_r$ shear catalogs. Each shape noise realization is generated with a different random seed. The same random seeds are used to generate shape noise catalogs across simulations with different cosmological parameters $\{\bb{p}_i\}$.

In addition to shape noise contributions to the observed galaxy ellipticity we consider photometric redshift errors as an additional contamination source in the simulated catalogs. In photometric surveys such as LSST, the source redshift $z_s$ is estimated measuring the source luminosity in a finite small set of optical frequency bands. Using this compressed luminosity information rather than the full spectrum introduces biases in redshift estimation. Forward modeling of the cosmic shear using the procedure described in \S~\ref{sec:shearsim}, as well as the shape noise contributions, assume a correct redshift distributions $n(z)$. An incorrect binning of observed galaxy redshifts according to the measured photometric distribution $n_p(z_p)$ can propagate the redshift measurement errors all the way to cosmological parameter constraints if the latter take advantage of redshift tomography. One of the goals of this work is to evaluate the size of this effect, assuming photometric redshift errors (photo-$z$) are left uncorrected. We model effect of photo-$z$ errors as a constant bias term $b_{\rm ph}(z_s)$ plus a random Gaussian component with variance $\sigma_{\rm ph}(z_s)$
\begin{equation}
\label{meth:photoz-correction}
z_p(z_s) = z_s + b_{\rm ph}(z_s) + \sigma_{\rm ph}(z_s)\mathcal{N}(0,1)   
\end{equation}
%
where $\mathcal{N}(0,1)$ is the standard normal distribution. We bin the $N_g$ galaxies in our simulated catalogs in 5 redshift bins $\bar{z}_b$, $b=1...5$. Several models have been proposed in the literature for the photometric bias $b_{\rm ph}(z_s)$ (see for example \citep{Huterer2006}) and variance $\sigma_{\rm ph}(z_s)$ (see for example \citep{LSSTSciBook}). We make use of the photo-$z$ bias and variance values contained in the LSST Science Book \citep{LSSTSciBook}, which we show in Figure \ref{fig:lsstscibook}. The bin sizes are adjusted so that each of them contains the same number of galaxies. The galaxy distribution we assumed, along with the binning choice is shown in Figure \ref{fig:galdistr}

\begin{figure}
\includegraphics[scale=0.3]{Figures/galdistr.eps}
\caption{Galaxy distribution assumed throughout this work (see equation (\ref{meth:galdistr})), along with the choice of the redshift bins $\bar{z}_b$. Our galaxy sample consists of $N_g=10^6$ galaxies. The figure shows the number of galaxies $N_g(z)$ at each redshift $z$.}
\label{fig:galdistr}
\end{figure}

@@

We measure a variety of summary statistics from the pixelized convergence maps, which will then be used to forecast parameter constraints and biases. We consider three kinds of summary statistics, namely the $\kappa$ tomographic power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, the tomographic peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and a set of $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$. The tomographic power spectrum is defined as 
\begin{equation}
\langle\tilde{\kappa}(\pmb{\ell},\bar{z}_b)\tilde{\kappa}(\pmb{\ell}',\bar{z}_{b'})\rangle = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})
\end{equation}
%
Because the $\kappa$ field is statistically isotropical, the expectation value $\langle\rangle$, for each realization $r$, is taken over all the modes $\pmb{\ell}$ which have the same magnitude $\ell=\vert\pmb{\ell}\vert$. The peak count statistic $n_{\rm pk}(\nu,\bar{z}_b)$ is defined as the number of the $\kappa$ local maxima of a certain height $\kappa_{\rm max}=\nu\sigma_0$, where $\sigma_0$ is the $\kappa$ standard deviation over pixels. The $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$ are defined as follows (see \citep{Matsubara10,Munshi12,MinkPetri}) 
%
\begin{equation}
\label{meth:momdef}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = (\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle) \\ \\
\pmb{\mu}_3 = (\langle\kappa^3\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle,\langle\kappa^2\nabla^2\kappa\rangle) \\ \\
\pmb{\mu}_4 = (\langle\kappa^4\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c)
\end{matrix}
\end{equation}
%
In equation (\ref{meth:momdef}) the gradients $\nabla$ are evaluated using finite differences between $\kappa$ values at neighboring pixels and the expectation values $\langle\rangle$ for each realization $r$ are taken over the $512^2$ pixels in the map. The subscript $c$ indicates that we consider only the connected parts of the quartic $\kappa$ moments. In the definition of the peak counts and convergence moments we omitted the redshift index $\bar{z}_b$ for notational simplicity, and also because both peak counts and moments do not contain cross redshift information unlike the power spectrum defined in equation (\ref{meth:psdefinition}). In the next paragraph we describe the statistical methods we use to infer cosmological parameter estimates $\bbh{p}$ from mock observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$ using the summary statistics $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, $n_{\rm pk}(\nu,\bar{z}_b)$, $\pmb{\mu}(\bar{z}_b)$.

@@

The $1\sigma$ square parameter errors $\Delta \bbh{p}^2$ are the diagonal entries of $\bbh{\Sigma}$. This method of forecasting parameters takes the name of Fisher Matrix formalism (see \citep{astroMLText} for a possible reference). When the dimension $N_b$ of the summary statistics space is large, numerical issues can arise in the estimation of the parameter error bars if the covariance matrix $\bbh{C}$ is measured from simulations. When $N_r$ independent realizations are used to estimate $\bbh{C}$, its inverse $\bbh{\Psi}$ is biased by a constant factor (see \citep{Hartlap07,Taylor12,Taylor14}) that can be taken into account. When the bias correction is applied, we can calculate the expectation value of the covariance estimator (\ref{meth:parcovestimator}) (see again\citep{Taylor14})
\label{meth:covdegradation}
\langle\bbh{\Sigma}\rangle = \left(\frac{N_r-N_b+N_p-1}{N_r-N_b-2}\right)\bb{\Sigma}
where $\bb{\Sigma}$ is the asymptotic covariance one obtains with an infinite number of realizations and $N_p$ is the number of parameters we are estimating, in this case 3. The scatter of the parameter estimates (\ref{meth:parestimate}) on the other hand scales as 
\label{meth:estdegradation}
\langle{\rm Cov}(\bbh{p})\rangle = \left(\frac{N_r-2}{N_r-N_b+N_p-2}\right)\bb{\Sigma}
Although equations (\ref{meth:covdegradation}) and (\ref{meth:estdegradation}) agree in the limit $N_r\rightarrow\infty$, they can be different when a finite number of realizations is used. More in detail, the degradation factor in the parameter covariance estimate in (\ref{meth:covdegradation}) is of order $1+(1+N_p)/N_r$, while the actual scatter of the estimates $\bbh{p}$ is of order $1+(N_b-N_p)/N_r$. These numbers can very different if $N_b$ is big. This means that the parameter error bar estimate (\ref{meth:covdegradation}) is too conservative if $N_b/N_r$ is of order unity. This could be likely the case in the analysis we are carrying because of the inclusion of tomography information. In order to avoid these error degradation issues, we apply dimensional reduction techniques to the summary statistics we are considering. These techniques are explained in the next paragraph.

>>>>

\title{Cosmology with Weak Lensing photometric surveys: constraints with redshift tomography of non--Gaussian statistics}

@@

\section{Introduction}
%
Weak Gravitational Lensing is a promising technique to probe the Large Scale Structure of the universe in which the tracers are unbiased \citep{wlreview}. This technique has the potential of improving significantly the constraints on the Dark Energy equation of state $w$ because it is most sensitive to the matter density fluctuations at the non--linear stage. Cosmology inferences from Weak Lensing observations have been produced for past (CFHTLenS \citep{cfht1}) and current (DES \citep{DES}) surveys, and are being planned for future experiments as well (LSST \citep{LSST}, WFIRST \citep{WFIRST}, Euclid \citep{Euclid}). Because of the non-linear nature of the density fluctuations probed by Weak Lensing, cosmological information might leak from quadratic statistics (such as two point functions and power spectra) into more complicated non-Gaussian statistics, for which forward modeling from cosmological parameters requires numerical simulations of cosmic shear fields. Different examples of these non-Gaussian statistics, and their cosmological information content, have been investigated extensively in the literature (see \citep{MinkJan,PeaksJan,NG-Marian,NG-Jain1,NG-Jain2,NG-Jain3,NG-Refregier,NG-Dietrich} for a non comprehensive list). The constraining power of Weak Lensing power spectra with the addition of redshift tomography information have been extensively investigated in the literature (see \citep{SongKnox,FangHaiman07,Huterer2006}). In this work we concentrate on the constraining power of a subset of these non-Gaussian statistics combined with redshift tomography in a LSST-like survey. Previous work on redshift tomography with Weak Lensing Minkowski functionals is also present in the literature \citep{MinkJan}. We also investigate the effects of uncorrected photometric redshift systematics on parameter constraints when using redshift tomography. This work is organized as follows: we give an outline of the shear simulations we use in this work, followed by descriptions of the convergence reconstruction procedure and forward modeling of galaxy shape and photometric redshift systematics. We then give an overview of the parameter inferences techniques we used to forecast constraints on cosmology. We then discuss the main results of this work and present our conclusions as well as prospects for future work.

@@

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating mock shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses until $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the lensing ODE up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from Large Scale Structure, assuming the non lensed shape is a circle. 
We simulate $N_g = 10^6$ random galaxy positions $\{\pmb{\theta}_g\}$ distributed uniformly in a field of view of size $\theta_{\rm FOV}^2=(3.5{\rm deg})^2$, which correspond to a galaxy angular density of $n_g=22\,\rm{arcmin}^{-2}$. The galaxies have a distribution in redshift which mimics the one expected in the LSST survey
\begin{equation}
\label{meth:galdistr}
n(z) = n_0\left(\frac{z}{z_0}\right)^2\exp\left(-\frac{z}{z_0}\right)
\end{equation}  
%
with $z_0=0.3$ and $n_0$ a normalization constant fixed so that $n(z)$ integrates to the total number of galaxies $N_g$. The galaxies have a maximum redshift $z_{\rm max}=3$. For each galaxy we compute the cosmic shear at $\pmb{\theta}_g$ using equation (\ref{meth:sheareqn}), producing a shear catalog $\{\pmb{\gamma}_g\}$. Different random realizations of a shear catalog $\{\pmb{\gamma}_g\}_r$ can be obtained rotating and periodically shifting the Large Scale Structure in the $N$--body snapshots according to the procedure explained in \citep{PetriVariance}. We produce $N_r=16000$ pseudo--independent realizations of the shear catalog $\{\pmb{\gamma}_g\}$. These shear realization cover the total survey area of LSST 10 times.
We repeat the above procedure for $P=100$ different combinations of the parameter triplet $\bb{p}$, sampled according to a Latin hypercube scheme. The sampling procedure is the same described in \citep{CFHTMink,CFHTPeaks}. The parameter space sampling we adopted for our simulations is shown in Figure \ref{fig:simdesign}

\begin{figure*}
\includegraphics[scale=0.35]{Figures/design.eps}
\caption{Parameter space sampling chosen for our simulations. We show both the sampling in the $(\Omega_m,w)$ (left), $(\Omega_m,\sigma_8)$ (center) and $(\sigma_8,w)$ (right) slices of the parameter space. The fiducial cosmology has been marked in red.}
\label{fig:simdesign}
\end{figure*}
%
For each of the parameter choices in Figure \ref{fig:simdesign}, the $N$--body initial conditions are generated using the same random seed. In addition to these simulations, we produce simulated shear catalogs for a fiducial $\Lambda$CDM universe with $\bb{p}_0=(0.26,-1,0.8)$. In this case the randomization procedure is based on 5 independent $N$--body simulations, and the same number $N_r=16000$ of pseudo--independent catalog realizations is produced. This additional simulation set serves two purposes: it provides an independent dataset from which to measure covariance matrices and it provides a way to construct mock observations that are independent on the simulations on which the cosmological forward model is trained. For the fiducial dataset we chose to base the shear randomization procedure on 5 independent $N$--body simulations to ensure the independence of the $N_r$ realizations for the purpose of estimating covariance matrices. \citep{PetriVariance} recently showed that, even with only one $N$--body simulation a few $10^4$ independent realizations can be produced.

@@

\subsection{Forward modeling of systematics}
We give an overview of the shear systematics included in this work.

The measured galaxy ellipticity $\pmb{\epsilon}$ is an estimate of the cosmic shear $\pmb{\gamma}$ due to Large Scale Structure if the non--lensed galaxy shape is a circle. If this assumption is relaxed, the measured galaxy ellipticity $\pmb{\epsilon}_{\rm m}$ can be modeled as a cosmic shear term plus a noise term \citep{wlreview}
\begin{equation}
\pmb{\epsilon}_{\rm m} = \pmb{\gamma} + \pmb{\epsilon}_{\rm n}
\end{equation} 
%
where $\pmb{\epsilon}_{\rm n}$ is a random Gaussian variable with zero mean and redshift dependent variance $\sigma_n(z)=0.15+0.035z$. This is equivalent to say that the cosmic shear inferred from ellipticity observations $\pmb{\gamma}_{\rm m}$ can be written as the sum of the true cosmic shear plus a noise term $\pmb{\gamma}_{\rm n}$ with the same statistical properties as $\pmb{\epsilon}_{\rm n}$. 
We add independent random realizations of the shape noise $\pmb{\gamma}_{\rm n}$ to each of the $N_r$ shear catalogs. Each shape noise realization is generated with a different random seed. The same random seeds are used to generate shape noise catalogs across simulations with different cosmological parameters $\{\bb{p}_i\}$.

In addition to shape noise contributions to the observed galaxy ellipticity we consider photometric redshift errors as an additional contamination source in the simulated catalogs. In photometric surveys such as LSST, the source redshift $z_s$ is estimated measuring the source luminosity in a finite small set of optical frequency bands. Using this compressed luminosity information rather than the full spectrum introduces biases in redshift estimation. Forward modeling of the cosmic shear using the procedure described in \S~\ref{sec:shearsim}, as well as the shape noise contributions, assume a correct redshift distributions $n(z)$. An incorrect binning of observed galaxy redshifts according to the measured photometric distribution $n_p(z_p)$ can propagate the redshift measurement errors all the way to cosmological parameter constraints if the latter take advantage of redshift tomography. One of the goals of this work is to evaluate the size of this effect, assuming photometric redshift errors (photo-$z$) are left uncorrected.

We model effect of photo-$z$ errors as a constant bias term $b_{\rm ph}(z_s)$ plus a random Gaussian component with variance $\sigma_{\rm ph}(z_s)$
\begin{equation}
\label{meth:photoz-correction}
z_p(z_s) = z_s + b_{\rm ph}(z_s) + \sigma_{\rm ph}(z_s)\mathcal{N}(0,1)   
\end{equation}
%
where $\mathcal{N}(0,1)$ is the standard normal distribution. We bin the $N_g$ galaxies in our simulated catalogs in 5 redshift bins $\bar{z}_b$, $b=1...5$. Several models have been proposed in the literature for the photometric bias $b_{\rm ph}(z_s)$ (see for example \citep{Huterer2006}) and variance $\sigma_{\rm ph}(z_s)$ (see for example \citep{LSSTSciBook}). We make use of the photo-$z$ bias and variance values contained in the LSST Science Book \citep{LSSTSciBook}, which we show in Figure \ref{fig:lsstscibook}. The bin sizes are adjusted so that each of them contains the same number of galaxies. The galaxy distribution we assumed, along with the binning choice is shown in Figure \ref{fig:galdistr}

\begin{figure}
\includegraphics[scale=0.3]{Figures/galdistr.pdf}
\caption{Galaxy distribution assumed throughout this work (see equation (\ref{meth:galdistr})), along with the choice of the redshift bins $\bar{z}_b$. Our galaxy sample consists of $N_g=10^6$ galaxies. The figure shows the number of galaxies $N_g(z)$ at each redshift $z$.}
\label{fig:galdistr}
\end{figure}

@@

We measure a variety of summary statistics from the pixelized convergence maps, which will then be used to forecast parameter constraints and biases. We consider three kinds of summary statistics, namely the $\kappa$ tomographic power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, the tomographic peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and a set of $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$. The tomographic power spectrum is defined as 
\begin{equation}
\langle\tilde{\kappa}(\pmb{\ell},\bar{z}_b)\tilde{\kappa}(\pmb{\ell}',\bar{z}_{b'})\rangle = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})
\end{equation}
%
Because the $\kappa$ field is statistically isotropical, the expectation value $\langle\rangle$, for each realization $r$, is taken over all the modes $\pmb{\ell}$ which have the same magnitude $\ell=\vert\pmb{\ell}\vert$. The peak count statistic $n_{\rm pk}(\nu,\bar{z}_b)$ is defined as the number of the $\kappa$ local maxima of a certain height $\kappa_{\rm max}=\nu\sigma_0$, where $\sigma_0$ is the $\kappa$ standard deviation over pixels. The $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$ are defined as follows (see \citep{Matsubara10,Munshi12,MinkPetri}) 
%
\begin{equation}
\label{meth:momdef}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = (\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle) \\ \\
\pmb{\mu}_3 = (\langle\kappa^3\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle,\langle\kappa^2\nabla^2\kappa\rangle) \\ \\
\pmb{\mu}_4 = (\langle\kappa^4\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c)
\end{matrix}
\end{equation}
%
In equation (\ref{meth:momdef}) the gradients $\nabla$ are evaluated using finite differences between $\kappa$ values at neighboring pixels and the expectation values $\langle\rangle$ for each realization $r$ are taken over the $512^2$ pixels in the map. The subscript $c$ indicates that we consider only the connected parts of the quartic $\kappa$ moments. In the definition of the peak counts and convergence moments we omitted the redshift index $\bar{z}_b$ for notational simplicity. In the next paragraph we describe the statistical methods we use to infer cosmological parameter estimates $\bbh{p}$ from mock observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$ using the summary statistics $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, $n_{\rm pk}(\nu,\bar{z}_b)$, $\pmb{\mu}(\bar{z}_b)$.

@@

The $1\sigma$ square parameter errors $\Delta \bbh{p}^2$ are the diagonal entries of $\bbh{\Sigma}$. The parameter covariance estimator (\ref{meth:parcovestimator}) is the same that one gets adopting a Fisher Matrix formalism for parameter forecasts (see \citep{astroMLText} for a possible reference).

When the dimension $N_b$ of the summary statistics space is large, numerical issues can arise in the estimation of the parameter error bars if the covariance matrix $\bbh{C}$ is measured from simulations. When $N_r$ independent realizations are used to estimate $\bbh{C}$, its inverse $\bbh{\Psi}$ is biased by a constant factor (see \citep{Hartlap07,Taylor12,Taylor14}) that can be taken into account. When the bias correction is applied, we can calculate the expectation value of the covariance estimator (\ref{meth:parcovestimator}) (see again\citep{Taylor14})
\begin{equation}
\label{meth:covdegradation}
\langle\bbh{\Sigma}\rangle = \left(\frac{N_r-N_b+N_p-1}{N_r-N_b-2}\right)\bb{\Sigma}
\end{equation}   
%
where $\bb{\Sigma}$ is the asymptotic covariance one obtains with an infinite number of realizations and $N_p$ is the number of parameters we are estimating, in this case 3. The scatter of the parameter estimates (\ref{meth:parestimate}) on the other hand scales as 
\begin{equation}
\label{meth:estdegradation}
\langle{\rm Cov}(\bbh{p})\rangle = \left(\frac{N_r-2}{N_r-N_b+N_p-2}\right)\bb{\Sigma}
\end{equation}
%
Although equations (\ref{meth:covdegradation}) and (\ref{meth:estdegradation}) agree in the limit $N_r\rightarrow\infty$, they can be different when a finite number of realizations is used. More in detail, the degradation factor in the parameter covariance estimate in (\ref{meth:covdegradation}) is of order $1+(1+N_p)/N_r$, while the actual scatter of the estimates $\bbh{p}$ is of order $1+(N_b-N_p)/N_r$. These numbers can very different if $N_b$ is big. This means that the parameter error bar estimate (\ref{meth:covdegradation}) is too conservative if $N_b/N_r$ is of order unity. This could be likely the case in the analysis we are carrying because of the inclusion of tomography information. In order to avoid these error degradation issues, we apply dimensional reduction techniques to the summary statistics we are considering. These techniques are explained in the next paragraph.
